---
title: "Fairfax Strategic Plan"
author: "Brandon L. Kramer"
date: "10/2/2019"
output: html_document
---

```{r setup, include = FALSE}
rm(list = ls())
knitr::opts_knit$set(root.dir = "~/Documents/Fairfax:INOVA/Fairfax Datasets")
for (pkg in c("tidyverse", "igraph", "bibliometrix", "tidytext", "ggplot2", "ggraph", "widyr", "plotly", "stringi", "pdftools", "SnowballC")) {library(pkg, character.only = TRUE)}
```

In this file, we are examining how [Fairfax County's Strategic Plans](https://www.fairfaxcounty.gov/strategicplan/sites/strategicplan/files/assets/documents/pdf/strategic-plan-preliminary-strategies.pdf). To get a start on understanding how topics (or in this case just words) are related between the various tenets (or EOs) of one section we ran some automated text analysis on this document. This process entailed pulling in the data, unnesting the tokens, removing the stopwords and then examining the bigram relationships between the words that most commonly occur within each of the tenets.   

```{r pulling data and basic word count, warning=FALSE, message=FALSE} 
# pulling the data
text_data <- read_csv("ffx_strategic_plan_text.csv") 
text_data <- as_tibble(text_data) 

# unnesting the tokens 
strategic_plan <- text_data %>% 
  group_by(section) %>% 
  unnest_tokens(word, text) %>%
  #mutate(word = wordStem(word)) %>%  # for stemming
  anti_join(stop_words)

# removing the stop words 
my_stopwords <- tibble(word = c(as.character(1:6), 
                                "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "18", "19", "22", 
                                "mt", "hnl", "eeg", "ell", "cro", "sspv", "ss", "eo"))
strategic_plan <- strategic_plan %>% anti_join(my_stopwords)

# getting the word counts 
word_counts <- strategic_plan %>%
  count(word, sort = TRUE)

# creating a dataframe of bigrams 
plan_bigrams <- strategic_plan  %>%
  unnest_tokens(bigram, word, token = "ngrams", n = 2)

# checking their counts 
plan_bigrams %>%
  count(bigram, sort = TRUE)

# converting the bigrams to edgelist format 
bigrams_separated <- plan_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")
bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)
```

```{r text networks, fig.width=9, fig.height=7}
set.seed(1234)
bigram_counts %>%
  filter(n >= 2) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "#F84C1E") +    #232D4B
  geom_node_point(size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 5,
                 point.padding = unit(0.4, "lines")) + theme_void()
```

```{r}
# and c
word_counts %>% 
  filter(word == "equity" | word == "equitable" | 
         word == "inequity" | word == "equality" | word == "inequality" |
         word == "diversity" | 
         word == "inclusion" | word == "inclusivity" | 
         word == "disparity" | word == "disparities")
```














