---
title: "Fairfax EO-Indicator Network"
author: "Brandon L. Kramer (UVA Biocomplexity Institute, SDAD)"
date: "10/16/2019"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include = FALSE}
rm(list = ls())

# load packages 
library(tidyverse)
library(networkD3)
library(igraph)

# pull the data 
# data available at: https://docs.google.com/spreadsheets/d/1025oveOpyS6GZNU--s7_HPITXGe5rN38akOXsRLi0Ac/edit#gid=0
setwd("~/Documents/FairfaxINOVA/Strategic Plan Networks/data")
edgeList <- read_csv("fairfax_eo-indicator_network-data - edgelist.csv")
nodeListAttributes <- read_csv("fairfax_eo-indicator_network-data - nodelist.csv")
```

```{r make network, echo=FALSE}
# this creates a graph and removes duplicated edges and self-loops 
gD <- igraph::simplify(igraph::graph.data.frame(edgeList, directed=FALSE))

# after creating that network, we need to construct a nodelist (starting at 0 because networkD3 is like that)
nodeList <- data.frame(ID = c(0:(igraph::vcount(gD) - 1)), nName = igraph::V(gD)$name)

# then we can assign the node names from the edge list to the nodelist 
getNodeID <- function(x){
  which(x == igraph::V(gD)$name) - 1 # to ensure that IDs start at 0
}

# and finally add them to them back to edgelist
edgeList <- plyr::ddply(edgeList, .variables = c("eo", "indicator", "weight"), 
                        function (x) data.frame(eoID = getNodeID(x$eo), 
                                                indicatorID = getNodeID(x$indicator)))

# next, we will calculate the degree centrality for all the nodes (and multiply them by 25 for visualization purposes)
nodeList <- cbind(nodeList, nodeDegree=15*igraph::degree(gD, v = igraph::V(gD), mode = "all"))

# we will also calculate the betweenness centrality for all nodes (and multiply them by 100 for visualization purposes)
betAll <- igraph::betweenness(gD, v = igraph::V(gD), directed = FALSE) / (((igraph::vcount(gD) - 1) * (igraph::vcount(gD)-2)) / 2)
betAll.norm <- (betAll - min(betAll))/(max(betAll) - min(betAll))
nodeList <- cbind(nodeList, nodeBetweenness=200*betAll.norm) 

# we also want the dice similarities between all pairs of nodes (although we won't use that in the visualizations below)
dsAll <- igraph::similarity.dice(gD, vids = igraph::V(gD), mode = "all")
F1 <- function(x) {data.frame(diceSim = dsAll[x$eoID +1, x$indicatorID + 1])}
edgeList <- plyr::ddply(edgeList, .variables=c("eo", "indicator", "weight", "eoID", "indicatorID"), 
                           function(x) data.frame(F1(x)))

# calculating group membership using cluster_walktrap in igraph  
wc <- cluster_walktrap(gD)
members <- membership(wc)
clusters <- igraph_to_networkD3(gD, group = members)
nodeList <- cbind(nodeList, cluster=clusters$nodes$group) 

# here, we want to add attributes to the nodeList to visualize different dataSources
nodeAttributes <- nodeListAttributes %>%
  select(name, label, dataSource) %>% 
  rename(nName = name)
nodeList <- plyr::join(nodeList, nodeAttributes, by="nName")
nodeList <- nodeList %>% remove_rownames %>% column_to_rownames(var="nName")

# and clean things up a little after we are done 
rm(betAll, betAll.norm, dsAll, F1, getNodeID, gD, nodeListAttributes, clusters)
```

## The Relations Between Strategic Plan EOs, Indicators and Datasources

**Visualization Note** First off, in order to see the interactive network, you will need to scroll on the visualization a little bit to center it within the frame. 

**Nodes** Now that you have it centered, you need to know a little about network analysis. All networks are composed of two core components: *nodes and ties*. In this network, the *nodes* correspond to two different entities. The deep red colored nodes are tenets were written in Fairfax County's Strategic Plan to promote the growth of economic opportunity. For short, we call these nodes EOs. The second type of node corresponds to indicators that have been used by other research institutions like [Broadstreet](https://www.broadstreet.io/board/BOARD/collection/Qm9hcmROb2RlOjI0ODE2/Food%20Access%20Report) to examine general trends about the population and resources in the county. 

**Colors** The bluish and greenish colored nodes are all indicators, but they are different colors because they refer to different data types. While the legend tells you the specifics, we wanted to simplify things since there are so many datasources. Dark blue nodes are datasets managed by the US government. The darker teal nodes are datasets managed by the state of Virginia or Fairfax County and the lighter turqoise nodes are datasources managed by private organizations. 

**Ties** When constructing this network, we wanted to know which indicators connect to each of the EOs. Below, the *ties* between the nodes show which indicators can help measure or better understand who could be affected by EO it shares a relation. As you play around with the visualization, there are a couple things that might help you orient this rather complex visualization. 

### Initial Results 

First, I would suggest mousing over all of the EOs to have some basis for what EOs are advocating. Once you have some familiarity, the next step is to hover over the indicators (in blue and green). You can that some indicators are really well connected, which you can more easily see by the *size* of the node. The most interesting part of the network is the collection of blue and green dots densely connected at the center of the graph. There, we can see that indicators like "Person Who Obtain Skills," "Persons With Improved Language Skills," and several demographic indicators all share connections to several EOs. These findings led us to some other questions (see below).

```{r visualizing our first network, echo=FALSE}
# creating a color palette for data types 
my_colors <- 'd3.scaleOrdinal() .domain(["Economic Opportunity Outcome (from Strategic Plan)", 
                                         
                                         "U.S. Census Bureau - Economic Census",
                                         "U.S. Bureau of Labor Statistics", 
                                         "U.S. Department of Agriculture - Food Environment Atlas",
                                         "U.S. Census Bureau - County Business Patterns",
                                         "Small Area Health Insurance Estimates", 
                                         "American Community Survey",
                                         "Feeding America",
                                         "National Center for Education Statistics",
                                         "Small Area Health Insurance Estimates",
                                         "U.S. Bureau of Labor Statistics",
                                         "U.S. Census Bureau",
                                         
                                         
                                         "Virginia Behavioral Risk Factor Surveillance System",
                                         "Virginia Department of Education",
                                         "Fairfax County Health and Human Services - Economic Self-Sufficiency",
                                         "Fairfax County Health and Human Services - Connected Individuals",
                                         "Fairfax County Health and Human Services - Healthy People",
                                         "Fairfax County Health and Human Services - Positive Living",
                                         "Fairfax County Health and Human Services - Sustainable Housing",
                                         
                                          
                                         "County Health Rankings", "United For ALICE"]) 
                                         
                                 .range(["#881234",
                                         "#065382", "#065382", "#065382", 
                                         "#065382", "#065382", "#065382",
                                         "#065382", "#065382", "#065382", 
                                         "#128177", "#128177", "#128177" , 
                                         "#128177", "#128177", "#128177", "#128177",  
                                         "#52BF90", "#52BF90"])'

# visualizing the network 
fn <- networkD3::forceNetwork(Links = edgeList, # data frame that contains info about edges
                        Nodes = nodeList, # data frame that contains info about nodes
                        Source = "eoID", # ID of source node 
                        Target = "indicatorID", # ID of target node
                        Value = "weight", # value from the edgelist that will be used to value/weight relationship amongst nodes
                        NodeID = "label", # value from the nodelist that contains node description we want to use (e.g., node name)
                        Nodesize = "nodeDegree",  # value from the nodelist that contains value we want to use for a node size
                        Group = "dataSource",  # value from the nodelist that contains value we want to use for node color
                        opacity = 0.9, # opacity
                        opacityNoHover = 0.2, # opacity of labels when static
                        height = 1800, # Size of the plot (vertical)
                        width = 2000,  # Size of the plot (horizontal)
                        fontSize = 45, # Font size
                        fontFamily = "sans-serif",
                        linkDistance = networkD3::JS("function(d) { return 125*d.value; }"),
                        zoom = TRUE, # ability to zoom when click on the node
                        charge = -300, # repulsion charge
                        colourScale = my_colors, # the color palette created above 
                        legend = TRUE) # including a legend 

htmlwidgets::onRender(
  fn, 'function(el, x) { 
    d3.selectAll(".legend")
    .style("font-size", "27px")
  }'
)
```

### The Importance of Fairfax Datasources

The network visualized below is composed of the same nodes (EOs and indicators) and ties as the graph above. The only difference is that we have highlighted the Fairfax County datasets to show how central they are to pursuing a data-driven approach to EO implementation. In particular, Fairfax County's Health and Human Services data seems to offer a number of informative indicators that could us develop a composite index (or potentially policies) that address more than one EO at a time. 

```{r highlighting ffx data in the network, echo=FALSE}
# creating a color palette for data types 
my_colors <- 'd3.scaleOrdinal() .domain(["Economic Opportunity Outcome (from Strategic Plan)", 
                                         
                                         "U.S. Census Bureau - Economic Census",
                                         "U.S. Bureau of Labor Statistics", 
                                         "U.S. Department of Agriculture - Food Environment Atlas",
                                         "U.S. Census Bureau - County Business Patterns",
                                         "Small Area Health Insurance Estimates", 
                                         "American Community Survey",
                                         "Feeding America",
                                         "National Center for Education Statistics",
                                         "Small Area Health Insurance Estimates",
                                         "U.S. Bureau of Labor Statistics",
                                         "U.S. Census Bureau",
                                         
                                         
                                         "Virginia Behavioral Risk Factor Surveillance System",
                                         "Virginia Department of Education",
                                         "Fairfax County Health and Human Services - Economic Self-Sufficiency",
                                         "Fairfax County Health and Human Services - Connected Individuals",
                                         "Fairfax County Health and Human Services - Healthy People",
                                         "Fairfax County Health and Human Services - Positive Living",
                                         "Fairfax County Health and Human Services - Sustainable Housing",
                                         
                                          
                                         "County Health Rankings", "United For ALICE"]) 
                                         
                                 .range(["#881234",
                                         "#065382", "#065382", "#065382", 
                                         "#065382", "#065382", "#065382",
                                         "#065382", "#065382", "#065382", 
                                         "#128177", "#128177", "#ff4c00" , 
                                         "#ff4c00", "#ff4c00", "#ff4c00", "#ff4c00",  
                                         "#52BF90", "#52BF90"])'

# visualizing the network 
fn <- networkD3::forceNetwork(Links = edgeList, # data frame that contains info about edges
                        Nodes = nodeList, # data frame that contains info about nodes
                        Source = "eoID", # ID of source node 
                        Target = "indicatorID", # ID of target node
                        Value = "weight", # value from the edgelist that will be used to value/weight relationship amongst nodes
                        NodeID = "label", # value from the nodelist that contains node description we want to use (e.g., node name)
                        Nodesize = "nodeDegree",  # value from the nodelist that contains value we want to use for a node size
                        Group = "dataSource",  # value from the nodelist that contains value we want to use for node color
                        opacity = 0.9, # opacity
                        opacityNoHover = 0.2, # opacity of labels when static
                        height = 1800, # Size of the plot (vertical)
                        width = 2000,  # Size of the plot (horizontal)
                        fontSize = 45, # Font size
                        fontFamily = "sans-serif",
                        linkDistance = networkD3::JS("function(d) { return 125*d.value; }"),
                        zoom = TRUE, # ability to zoom when click on the node
                        charge = -300, # repulsion charge
                        colourScale = my_colors, # the color palette created above 
                        legend = TRUE) # including a legend 

htmlwidgets::onRender(
  fn,
  'function(el, x) { 
    d3.selectAll(".legend")
    .style("font-size", "27px")
    .style("margin", "1000px 1000px 1000px 1000px")
  }'
)
```

### Using Network Analysis to Detect Indicator Clusters for Composite Indices

Speaking of composite indices, there are also network statistics that can help us anlayze different clusters (or communities) of EOs and indicators. Using a community detection algorithm that calculates clusters based on a measure of similarity between nodes that is based on random walks throughout the network (see Pons and Latapy 2006). This analysis shows that there are five larger clusters of indicators that can help us potentially develop composite indices for our CommunityScapes that feed more directly into addressing the EOs. Take a look at these clusters below. 

```{r community detection, echo=FALSE}
# creating a color palette for data types 
cluster_colors <- 'd3.scaleOrdinal() .domain(["1", "2", "3", "4", "5"])
                                         
                                 .range(["#881234", "#065382", "#128177", "#52BF90", "#52BF90"])'

# visualizing the network 
fn <- networkD3::forceNetwork(Links = edgeList, # data frame that contains info about edges
                        Nodes = nodeList, # data frame that contains info about nodes
                        Source = "eoID", # ID of source node 
                        Target = "indicatorID", # ID of target node
                        Value = "weight", # value from the edgelist that will be used to value/weight relationship amongst nodes
                        NodeID = "label", # value from the nodelist that contains node description we want to use (e.g., node name)
                        Nodesize = "nodeDegree",  # value from the nodelist that contains value we want to use for a node size
                        Group = "cluster",  # value from the nodelist that contains value we want to use for node color
                        opacity = 0.9, # opacity
                        opacityNoHover = 0.2, # opacity of labels when static
                        height = 1800, # Size of the plot (vertical)
                        width = 2000,  # Size of the plot (horizontal)
                        fontSize = 45, # Font size
                        fontFamily = "sans-serif",
                        linkDistance = networkD3::JS("function(d) { return 125*d.value; }"),
                        zoom = TRUE, # ability to zoom when click on the node
                        charge = -300, # repulsion charge
                        colourScale = cluster_colors, # the color palette created above 
                        legend = TRUE) # including a legend 

htmlwidgets::onRender(
  fn, 'function(el, x) { 
    d3.selectAll(".legend")
    .style("font-size", "27px")
  }'
)
```

## Citations 

https://www.r-bloggers.com/network-visualization-part-6-d3-and-r-networkd3/
https://stackoverflow.com/questions/57396295/add-an-vertex-attribute-to-networkd3-object
http://emis.ams.org/journals/JGAA/accepted/2006/PonsLatapy2006.10.2.pdf














